### 2025-09-11 docs/test.tex
old latex is corruped, work on this one in the meantime

This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers. A growing literature emphasizes that families matter for economic outcomes \citep{Doepke2016a}. In the context of homeownership, previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}. Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}. However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not address the overall importance of parental transfers. For example, households may anticipate future transfers and subsequently save less. In this case, even if transfers increase the likelihood of homeownership, all else being equal, they may not affect aggregate homeownership. Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.

I first present empirical evidence showing that parental wealth helps not only with the transition into ownership, but also in retention of ownership. In addition to replicating previous findings that transfers increase the probability of becoming a homeowner and that parental wealth is associated with sustained ownership \citep[see e.g.,][]{Lee2018,bond2021role}, I provide novel results on the mechanism through which parental wealth aids retention. Controlling for a wide set of household and parental characteristics, I show that households with wealthier parents are less likely to fall behind on mortgage payments. Finally, using a variation of the event study in \cite{Chetty2007}, I show that while households with less wealthy parents downsize their housing during unemployment spells, those with wealthy parents retain their current homes.

@@

I then examine the interplay  of transfers and housing for marginal propensities to consume (MPC). A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.  In my model, the introduction of altruistic transfers lowers average MPCs by about 20\%, implying that standard life-cycle models may understate the discrepancy.  Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 30\%. Another explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---yet such behavior is observationally similar to that of households with wealthy, altruistic parents.

@@

	\caption*{\begin{footnotesize} \textit{Notes:}  Left panel (SHED): sample is current owners reporting this is their first home; x-axis is purchase year; outcome is the share receiving any down-payment transfer. Right panel (AHS): sample is owner-occupied units that do not report proceeds from a prior home sale as the main funding source; x-axis is survey year; outcome is the share whose down payment was mostly funded by a transfer. Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
\end{figure}

@@

In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960. Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own. Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy. The key determinant of transfer receipt seems to be parental wealth.

@@

Next, I break the sample down by ownership. Homeowners are both wealthier and have wealthier parents than renters. Renters and owners receive transfers at about the same rate and size. Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.

@@

To understand which variables predicts retention of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where $Own_{i,t+\Delta}=1$ denotes ownership at time $t+\Delta$, $\beta$, and $\zeta$ measure the effects of log household net worth and parental wealth, respectively. The vector of control variables  $X_{i,t}$ consists of log household income, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.  The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations using the PSID and the HRS. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plot the two coefficients of interest over different time horizons.

@@

I first limit the sample to first-time buyers to focus on distress among new owners. In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$). With the full set of controls, the effect increases slightly, though it is no longer statistically significant. These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage. In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects. In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed-effects regression ($p<0.1$).

>>>>

This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers. A growing literature emphasizes that families matter for economic outcomes \citep{Doepke2016a}. In the context of homeownership, previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}. Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}. However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not establish the overall importance of parental transfers. Households may anticipate future transfers and reduce their own saving, so that the observed boost in homeownership conditional on receiving a transfer overstates the aggregate effect of transfers on homeownership.Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.

I first present empirical evidence showing that parental wealth helps not only with the transition into ownership, but also in retention of ownership. In addition to replicating previous findings that transfers increase the probability of becoming a homeowner and that parental wealth is associated with sustained ownership \citep[see e.g.,][]{Lee2018,bond2021role}, I provide novel evidence on one mechanism: households with wealthier parents are less likely to fall behind on mortgage payments, consistent with parental resources providing insurance against income shocks. Finally, using a variation of the event study in \cite{Chetty2007}, I show that while households with less wealthy parents downsize their housing during unemployment spells, those with wealthy parents retain their current homes.

@@

Finally, I examine the interplay  of transfers and housing for marginal propensities to consume (MPC). A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.  In my model, the introduction of altruistic transfers lowers average MPCs by about 20\%, implying that standard life-cycle models may understate the discrepancy.  Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 30\%. Another explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---yet such behavior is observationally similar to that of households with wealthy, altruistic parents.

@@

	\caption*{\begin{footnotesize} \textit{Notes:}  Left panel (SHED): first-time owners by purchase year; outcome is the share receiving a down-payment transfer. Right panel (AHS): owner-occupied units without prior-sale proceeds by survey year; outcome is the share whose down payment was primarily funded by a transfer. Higher levels in the SHED reflect its broader definition. Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
\end{figure}

@@

In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960. Receivers have parents about 2.5 times richer, have similar wealth and income as nonreceivers, and are less likely to own. Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy. The key determinant of transfer receipt seems to be parental wealth.

@@

Next, I split the sample by ownership. Homeowners are wealthier and have wealthier parents than renters. Renters and owners receive transfers at about the same rate and size. Among current owners, 21\% of receivers rented two years earlier, compared with 14\% of nonreceivers, suggesting that transfers ease the transition into ownership.

@@

To understand which variables predict retention of ownership, I estimate
\begin{equation}	\label{eq:regmaintain}
	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
	  + \gamma' X_{i,t} + \varepsilon_{i,t},
\end{equation}
where $Own_{i,t+\Delta}=1$ denotes ownership at time $t+\Delta$, $\beta$, and $\zeta$ measure the effects of log household net worth and parental wealth, respectively. The vector of control variables  $X_{i,t}$ consists of log household income, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.  The sample is limited to households aged 25 to 44. These regressions are motivated by \cite{bond2021role}, who estimate similar equations using the PSID and the HRS. Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plot the two coefficients of interest over different time horizons.

@@

I first limit the sample to first-time buyers to focus on distress among new owners. In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$). With the full set of controls, the effect increases slightly, though it is no longer statistically significant. These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage and parents of receivers are about 2.5 richer on average. In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects. In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed-effects regression ($p<0.1$).
